---
title: "Lab1"
output: html_document
---

Your Turn: Play with the bands
======================================

Here is a snippet of code that creates a list containing information about three
rock bands. 

```{r}
options(stringsAsFactors = FALSE)
bands <- list(list(name = "The Jimi Hendrix Experience",
                   years = 1966:1969,
                   members = data.frame(name = c("Jimi Hendrix", "Noel Redding", "Mitch Mitchell"),
                                        instrument = c("guitar", "bass", "drums"))),
              list(name = "Nirvana",
                   years = 1987:1994,
                   members = data.frame(name = c("Kurt Cobain", "Krist Novoselic", "Dave Grohl"),
                                        instrument = c("guitar", "bass", "drums"))),
              list(name = "The White Stripes",
                   years = 1997:2011,
                   members = data.frame(name = c("Jack White", "Meg White"),
                                        instrument = c("guitar", "oreos"))))
```

1. Extract the `years` element of the second band in the list.
   * What class is it (integer, numeric, logical ...)?
   * How many years was the band active?
2. Extract the name of the bass player in the second band.
3. Meg White played drums in the band "The White Stripes" (the third band in the
   list). Correct the data to reflect this fact.
4. (Advanced) Combine all three `members` elements into a single data.frame.
5. (Advanced) Sort the combined data.frame by `instrument`. Break any ties by
   sorting by `name`.
   
```{r}
# 1. extract `years` from second band
second_band <- bands[[2]]
print(second_band$years)
print(class(second_band$years))
print(length(second_band$years))

# 2. extract the name of the bass player in the second band
second_band_members <- second_band$members
print(second_band_members$name[second_band_members$instrument == "bass"])

# 3. Correct Meg White info in the third band
bands[[3]]$members$instrument[bands[[3]]$members$name == "Meg White"] <- "drums"
print(bands[[3]])

# 4. Combine all three members elements into a single data frame
all_members <- rbind(bands[[1]]$members, bands[[2]]$members, bands[[3]]$members)
print(all_members)

# 5. Sort members by instrument
sorted_members <- all_members[order(all_members$instrument, all_members$name), ]
print(sorted_members)
```

Your Turn: Read and inspect data
================================

1. Using R, create a directory named `lab0data`.
2. Using R, download the file at
   `http://tutorials.iq.harvard.edu/example_data/hp1602.csv` to your
   `lab0data` directory. Name the file "housingPriceIndex2016.csv". 
3. Read the data into R, assigning the result to the name `housing`.
4. (Advanced) The `.csv` file above was extracted from the Excel file at
   `https://www.dallasfed.org/institute/houseprice/~/media/documents/institute/houseprice/hp1602.xlsx`
   Use R to download this file.
5. (Advanced) Open the file downloaded in step 4 in a spreadsheet application
   and determine what needs to be done to read it into R. For example, which
   sheet should be read? Do you need to skip some rows?
6. (Advanced) Read the data downloaded in step 4 into R. 

```{r}
setwd("~/HARVARD/Courses/18S_CS109b/2018-cs109b/Labs/Lab1")
dir.create("lab0data")
download.file(url = "http://tutorials.iq.harvard.edu/example_data/hp1602.csv", destfile = "lab0data/housingPriceIndex2016.csv", method = "curl")
housing = read.csv("lab0data/housingPriceIndex2016.csv")


```